---
title: "Dashboard_Centenarians"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
runtime: shiny
---

```{r setup, include=FALSE}
library(flexdashboard)
library(DT)
library(shiny)
```


```{r global, include=FALSE}
library(here)
library(tidyverse)

if (!dir.exists(here::here("data"))) {
    dir.create(here::here("data"))
}

# saves data only once (not each time you knit a R Markdown)
if (!file.exists(here::here("data", "centenarians.rda"))) {
     tuesdata <- tidytuesdayR::tt_load("2023-05-30")
     centenarians <- tuesdata$centenarians
     save(centenarians, file = here::here("data", "centenarians.rda"))
 }

#Read in the data locally each time you knit/render.
load(here::here("data", "centenarians.rda"))

centenarians <- centenarians %>% mutate(
  gender = factor(gender),
  still_alive = factor(still_alive)
)
```


# About

## Column 1 {data-width=350}

### **Purpose of this dashboard**

This dashboard is for designed for everyone who are interested in centenarians around the world. The purpose of this dashboard is to provide the audience with detailed information on these centenarians, such as their name, their country of residence, their sex, and their birth date. It also guides audience in exploring differences in life expectancy with respect to characteristics such as country of residence and sex. 

The second purpose of this dashboard is to practice building dashboards with R.


### **Source of the original dataset**

I used the [Centenarian](https://github.com/rfordatascience/tidytuesday/blob/master/data/2023/2023-05-30/centenarians.csv) data from Tidytuesday. 



## Column 2

### **A video explaining this dashboard**

```{r}
youtube_url <- "https://www.youtube.com/embed/9EXLUDKbXxY"
iframe_code <- tags$iframe(src = youtube_url, width = "700", height = "450", frameborder = "0", allowfullscreen = TRUE)
iframe_code
```


### New Updates on Changes in my Dashboard {data-height=200}

In the "Top 3 Longest Lived Centenarians" page, the picture of these people do not show up after deploying the dashboard, even though I used here() function, it still did not work. I included a histogram of their ages instead of including pictures of them. 



# Data


**About the Data**

This [dataset](https://github.com/rfordatascience/tidytuesday/blob/master/data/2023/2023-05-30/centenarians.csv) contains 100 female centenarians and 100 male cenetarians all around the world who are in records. This dataset contains their name, country of residency, age, birth date, death date (if exist), sex, and whether they are still alive. For detail information, please refer to the [data dictionary here.](https://github.com/rfordatascience/tidytuesday/tree/master/data/2023/2023-05-30)



```{r DTprep, echo=FALSE}
# Create the dataframe for the DT table
DT_df <- centenarians 
```

```{r DT, echo=FALSE}
# Make the DT table
DT::renderDataTable({
  DT::datatable(DT_df,
                options = list(autoWidth = TRUE,
                               pageLength = 10,
                               scroller = TRUE,
                               scrollY = '100px'))
})

```





# World Centenarians 


## Column {.sidebar}

```{r}
selectizeInput("country", "Select Country", choices = sort(unique(centenarians$place_of_death_or_residence)), selected = c("United States", "United Kingdom", "Japan", "Spain", "France"), multiple = TRUE)
```

## Output 

```{r}
filtered_data <- reactive({
  centenarians[centenarians$place_of_death_or_residence %in% input$country, ] %>% 
    na.omit() %>%
    group_by(place_of_death_or_residence) %>% 
    summarise(n = n())
})

renderPlot({
  ggplot(filtered_data(), aes(x = "", y = n, fill = place_of_death_or_residence)) +
    geom_bar(stat = "identity", width = 1, color = "white") +
    coord_polar(theta = "y") +
    scale_fill_manual(values = rainbow(length(unique(filtered_data()$place_of_death_or_residence)))) +
    theme_minimal() +
    labs(title = "Percentage of Selected Countries")
})

```


# Centenarians File

## Column {.sidebar}




```{r}

selectizeInput("gender", "Select Gender", unique(centenarians$gender))
selectizeInput("status", "Select Status", unique(centenarians$still_alive))

sliderInput("age", "Select Age Range", 
            min = round(min(centenarians$age), 5), 
            max = round(max(centenarians$age), 5),
            value = c(min(centenarians$age), max(centenarians$age)))
```


## Output 

```{r}
filter_data <- reactive({
  filtered <- centenarians %>%
    filter(age >= input$age[1] & age <= input$age[2],
           gender %in% input$gender,
           still_alive %in% input$status)

  return(filtered)
})

renderTable({
  filter_data()
})

```


# Centenarians' Birthdays


## column {.tabset}


### Deceased

```{r}
centenarians_deceased <- centenarians %>% filter(still_alive == "deceased")
ggplot(centenarians_deceased) + geom_histogram(aes(x = birth_date), fill = "blue", bins = 50) + theme_bw() + labs(title = "Histogram of World Centenarians' Birthdays", x = "Birth Date", subtitle = "Most of the centenarians were born after 1900.", caption = "Meilin Zheng") + scale_x_date(date_breaks = "10 years")
```


### Alive

```{r}
centenarians_alive <- centenarians %>% filter(still_alive == "alive")
ggplot(centenarians_alive) + geom_histogram(aes(x = birth_date), fill = "purple") + theme_bw() + labs(title = "Histogram of World Centenarians' Birthdays", x = "Birth Date", subtitle = "The latest-borned centenarian was born in 1912.", caption = "Meilin Zheng")
```


# Top 3 Longest Lived Centenarians


## Column {.tabset}

### Female 

**Jeanne Calment**

**Kane Tanaka**

**Sarah Knauss**

```{r}
female <- centenarians %>% filter(name == "Jeanne Calment" | name == "Kane Tanaka" | name == "Sarah Knauss")
ggplot(female) + geom_histogram(aes(x = age), bins = 50) + labs(title = "Top 3 Logest-lived Female Centenarians", caption = "Meilin Zheng")
```


### Male

**Antonio de Castro**

**Francisco Juárez Iglesias**

**Reuben Sinclair**

```{r}
male <- centenarians %>% filter(name == "Antonio de Castro" | name == "Francisco Juárez Iglesias" | name == "Reuben Sinclair")
ggplot(male) + geom_histogram(aes(x = age), bins = 50) + labs(title = "Top 3 Logest-lived Male Centenarians", caption = "Meilin Zheng")
```


# Analysis

## Column


**Question: Are there any associations between life expectancy, gender, country of residence, and birth date?**


```{r wrangle, message = FALSE, warning = FALSE}
analysisdata <- centenarians

analysisdata <- analysisdata %>% mutate(
  country = place_of_death_or_residence,
  sex = gender, 
  status = still_alive)

```


**Life Expectancy among Males and Females**

```{r analysis,  message = FALSE, warning = FALSE}
table_mf <- analysisdata %>% group_by(sex) %>% summarise(mean_life_expectancy = sum(age)/n()) 

renderTable({
  table_mf
})
#used functions
#group_by(), summarise(), n()
```


The life expectancy among oldest females is higher than the life expectancy among oldest males.


```{r, fig.height=5, fig.width=7.5}
age_sex <- analysisdata %>% group_by(sex) %>% summarise(mean_life_expectancy = sum(age)/n()) 
age_sex <- data.frame(age_sex)

library(ggplot2)
p <- ggplot() + geom_boxplot(data = analysisdata, aes(x = sex, y = age)) + geom_point(data = age_sex, aes(y = mean_life_expectancy, x = sex), shape = 18, color = "red", size = 6) + labs(title = "Life Expectancy by Sex", subtitle = "The red diamond is the calculated mean life expectancy among these two group. We can see that\nwomen has a higher mean life expectancy than men. Also, in the upper 25 percentile, women also\n have a much higher life expectancy than men.", y = "Life Expectancy", caption = "Meilin Zheng") + theme_bw()

#function used 
#geom_boxplot(), geom_point()

p

```



**Life Expectancy by Country**

```{r}
table_country <- analysisdata %>% group_by(country) %>% summarise(total = n()) %>% mutate(country_lump = fct_lump(country, 6)) %>% group_by(country_lump) %>% summarise(total = sum(total)) %>% arrange(desc(total)) %>% head(6)


renderTable({
  table_country
})

#function used
#fct_lump()

```

The table above gives the t6 countries with highest number of oldest people.

In the table above, we can see that among the 200 oldest people, 66 of them are from the United States, and 46 of them are from Japan.

By observing this table, it is still questionable whether United States and Japan have the highest number of oldest people in the world. The dataset may have a sampling bias that did not got to include oldest adult from other places around the world.

However, we will still use these 6 countries for analysis since they have larger amount of samples.

```{r}
analysisdata_reduced <- analysisdata %>% filter(country == "United States" | country == "Japan" | country == "France" | country == "Italy" | country == "Spain" | country == "United Kingdom")
#functions used
#filter()
table_redcued <- analysisdata_reduced %>% group_by(country) %>% summarise(mean_life_expectancy = sum(age)/n())

renderTable({
  table_redcued
})
```


Among these six countries, people from France have the highest life expectancy.

```{r, fig.height=5, fig.width=7.5}
ggplot(analysisdata_reduced) + geom_histogram(aes(x = age)) + geom_vline(aes(xintercept = mean(age)), color = "red", linetype = "dashed", size = 1) + theme_bw() + facet_wrap(country~.) + geom_vline(data = (analysisdata_reduced %>% group_by(country) %>% summarise(mean_age = mean(age))), aes(xintercept = mean_age), color = "green", linetype = "dashed", size = 1) + labs(title = "Histogram of Life Expectancy Distribution by Countries", x = "Life Expectancy", subtitle = "The dashed red line is the grand mean among these 6 countries, whereaes the dashed green line is the\nmean life expectancy within each country. We can see that France has a higher mean life expectancy\nthan the overal mean, while the United States, Japan, and Italy have a similar mean. Spain and the\nUnited Kingdom have a slightly lower mean life expectancy than the grand mean.", caption = "Meilin Zheng")

#used function
#geom_histogram(), geom_vline(), facet.wrap()
```


**Life Expectancy by Birth Time**

```{r}
centenarians_year <- analysisdata %>% mutate(Year = year(birth_date))


centenarians_year_1900 <- centenarians_year %>% filter(Year < 1900)
p1 <- ggplot(centenarians_year_1900) + geom_point(aes(x = Year, y = age)) + theme_bw() + labs(title = "Birth Year vs. Life Expectancy\nbefore 20th Century", subtitle = "No obious association is observed between\nbirth time and age.", caption = "Meilin Zheng", y = "Life Expectancy") + scale_y_continuous(limits = c(110, 125))

p1


```

```{r}
centenarians_year_2000 <- centenarians_year %>% filter(Year > 1900)
p2 <- ggplot(centenarians_year_2000) + geom_point(aes(x = Year, y = age)) + theme_bw() + labs(title = "Birth Year vs. Life Expectancy\nafter 20th Century", subtitle = "No obious association is observed between\nbirth time and age.", caption = "Meilin Zheng", y = "Life Expectancy") + scale_y_continuous(limits = c(110, 125))

p2

```


**Summary**

We found that female have a higher life expectancy than male. Also, developed countries such as the United States and Japan seem to have more people with greater life expectancy. There is no obvious association observed between the time the people were born and their life expectancy.






